A note on $\K$ of Kac---Moody groups over a field

### 2017-08-24 paper.tex
Fixes as in proof

\subsection{The main result}
Now we turn to the case when $A$ contains columns of both types 
(i.\,e. there are columns having an odd entry, and consisting only of even entries).
The basic idea remains the same: we need to find a nice basis for the subgroup $J\leq L$ of defining relations of $\K(A, F)$.
This time, however, the situation is more tricky since we cannot simply tensor out $\K(F)$ or $\Kmw(F)$.
To solve the problem we need to invent some analogue of the fundamental theorem of finitely-generated abelian groups suitable to our sitation.

@@

Suppose that $i$-th column of $c$ is odd and $j$-th one is even.

We only need to consider the case $|c_{ki}| < |c_{kj}|$. Write $c_{kj} + q c_{ki} = r$ for some $q$ and $r$, satisfying $|r| < |c_{ki}|$.
If $q$ is even, add $q \cdot c_{*i}$ to $c_{*j}$ using an operation of type~\eqref{item:type5} and 
 notice that $\mathrm{min}(|c_{ki}|,|r|) < \mathrm{min}(|c_{ki}|,|c_{kj}|)$.
Otherwise, choose even $q'$ equal to either $q-1$ or $q+1$ so that one of the following is true:
\begin{itemize}
 \item $r\neq 0$ and $|c_{kj} + q' c_{ki}| < |c_{ki}|$;
 \item $r = 0$ and $c_{kj} + q' c_{ki} = c_{ki}$.
\end{itemize}
In the first case after adding $q' \cdot c_{*i}$ to $c_{*j}$ we obtain the required inequality.
In the second case we need to further subtract $c_{*j}$ from $c_{*i}$ to get $c_{ki} = 0$.
\end{proof}

@@

Looking at the effect of operations \eqref{item:type2}--\eqref{item:type5} modulo $2$ one can notice the following two properties of the matrix $q$:
\begin{itemize}
 \item $i$-th column of $d$ is odd if and only if coefficients $q_{ij}$ are even for $k+1\leq j \leq n$;
 \item $i$-th column of $d$ is odd if and only if one of the coefficients $q'_{ji}$ of the inverse matrix $q^{-1}$ is odd for some $1\leq j\leq k$.
\end{itemize}

In \cite{MW} M.~Westaway even conjectured that the above two properties are simultaneously fulfilled iff $p_d(i)=odd$, and thus, that
 it is possible to determine the type of summands appearing in $\K(A, F)$ from the given matrix $q$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correspond to our $q$ via $\nu = q^T$, $\mu = (q^{-1})^T$).
Although, as said before, this characterization works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for {\it arbitrary} $q$ occurring in SNF decomposition of $\widetilde{A}$
 (e.\,g. the one obtained via the classical algorithm for finding SNF, which ignores columns' parities).
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.

\begin{example}
To produce a counterexample, start with GCM 
 $A = \left(\begin{smallmatrix}
 2& -1& -2\\
-3& 2& -4\\
-4& -2& 2\end{smallmatrix}\right)$
  and write down SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$:
\[\widetilde{A} = \left(\begin{smallmatrix}
-3& 1& 0 \\
-4& 0& 2 \\
 0& -2& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
 1& 0& 0\\
 0& 1& 0\\
-2& 2& 1\end{smallmatrix}\right) \cdot \mathrm{diag}(1,2,2) \cdot \left(\begin{smallmatrix}
-3& 1& 0\\
-2& 0& 1\\
 1& 0& 0 \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last column of $q$ we get that $1$st and $3$rd columns of $d$ are odd while $2$nd one is even,
 hence $\K(A, F) \cong \Kmw(F)/I \oplus \K(F)/2\K(F)$.

>>>>

\subsection{The main result}
Now we turn to the case when $A$ contains columns of both types 
(i.\,e. there are columns having an odd entry, and consisting only of even entries).
The basic idea remains the same: we need to find a nice basis for the subgroup $J\leq L$ of defining relations of $\K(A, F)$.
This time, however, the situation is more tricky since we cannot simply tensor out $\K(F)$ or $\Kmw(F)$.
To solve the problem we need to invent some analogue of the fundamental theorem of finitely-generated abelian groups suitable to our situation.

@@

Suppose that $i$-th column of $c$ is odd and $j$-th one is even.
We only need to consider the case $|c_{ki}| < |c_{kj}|$. Write $c_{kj} + q c_{ki} = r$ for some $q$ and $r$, satisfying $|r| < |c_{ki}|$.
If $q$ is even, add $q \cdot c_{*i}$ to $c_{*j}$ using an operation of type~\eqref{item:type5} and 
 notice that $\mathrm{min}(|c_{ki}|,|r|) < \mathrm{min}(|c_{ki}|,|c_{kj}|)$.
Otherwise, choose even $q'$ equal to either $q-1$ or $q+1$ so that one of the following is true:
\begin{itemize}
 \item $r\neq 0$ and $|c_{kj} + q' c_{ki}| < |c_{ki}|$;
 \item $r = 0$ and $c_{kj} + q' c_{ki} = c_{ki}$.
\end{itemize}
In the first case after adding $q' \cdot c_{*i}$ to $c_{*j}$ we obtain the required inequality.
In the second case we need to further subtract $c_{*j}$ from $c_{*i}$ to get $c_{ki} = 0$.
\end{proof}

@@

Looking at the effect of operations \eqref{item:type2}--\eqref{item:type5} modulo $2$ one can notice the following two properties of the matrix $q$:
\begin{itemize}
 \item $i$-th column of $d$ is odd iff coefficients $q_{ij}$ are even for $k+1\leq j \leq n$;
 \item $i$-th column of $d$ is odd iff one of the coefficients $q'_{ji}$ of the inverse matrix $q^{-1}$ is odd for some $1\leq j\leq k$.
\end{itemize}

In \cite{MW} M.~Westaway even conjectured that $i$-th column of $d$ is odd iff the above two properties of $q$ are simultaneously fulfilled, and thus, that
 it is possible to determine the type of summands appearing in $\K(A, F)$ from the given matrix $q$
 (Westaway's matrices $\nu$, $\mu$ appearing in the statement of \cite[Conjecture~9.1]{MW} correspond to our $q$ via $\nu = q^T$, $\mu = (q^{-1})^T$).
Although, as said before, this characterization works for the matrix $q$ obtained via the algorithm of~\cref{lm:normal-form},
 it may not work for {\it arbitrary} $q$ occurring in SNF decomposition of $\widetilde{A}$
 (e.\,g. the one obtained via the classical algorithm for finding SNF, which ignores columns' parities).
Thus, as stated, \cite[Conjecture~9.1]{MW} is false.

\begin{example}
To produce a counterexample, start with GCM 
 $A = \left(\begin{smallmatrix}
 2& -1& -2\\
-3& 2& -4\\
-4& -2& 2\end{smallmatrix}\right)$
  and write down SNF decomposition of~\cref{lm:normal-form} for $\widetilde{A}$:
\[\widetilde{A} = \left(\begin{smallmatrix}
-3& 1& 0 \\
-4& 0& 2 \\
 0& -2& 4
\end{smallmatrix} \right) = \left(\begin{smallmatrix}
 1& 0& 0\\
 0& 1& 0\\
-2& 2& 1\end{smallmatrix}\right) \cdot \mathrm{diag}(1,2,2) \cdot \left(\begin{smallmatrix}
-3& 1& 0\\
-2& 0& 1\\
 1& 0& 0 \end{smallmatrix}\right) = p \cdot d \cdot q.\]
Looking at the last column of $q$ we conclude that $1$st and $3$rd columns of $d$ are odd while $2$nd one is even,
 hence $\K(A, F) \cong \Kmw(F)/I \oplus \K(F)/2\K(F)$.
